---
title: "ENVS 193DS Homework 3"
author: "Ellie Schomberg"
date: "05-28-2025"
format: html
---

Link to GitHub repository is [here](https://github.com/ellieschomberg/ENVS-193DS_homework-03.git)

# Set Up

```{r set up}
#| message: FALSE
# reading in packages
library(tidyverse)
library(here)
library(flextable)
library(janitor)
library(readxl)

# reading in data
personal_data <- read.csv("data/updated_personal_data.csv")
```

# Problem 1. Personal Data

## a. Summarizing data

For my project, I am testing to see if the type of swimming workout that I do, each with different yardages and levels of intensity, influences the average amount of sleep I get per night. I am comparing the mean between groups, and I believe that more intense workouts will cause me to sleep more.

## b. Visualization

```{r cleaning data}
personal_data_clean <- read.csv("data/updated_personal_data.csv") |> # start with updated_personal_data data frame
  clean_names() # clean up the column names
```

```{r summarizing data}
personal_data_summary <- personal_data_clean |> # start with personal_data_clean dataframe
  group_by(type_of_workout) |>  #grouping by type of workout
  summarize(mean = mean(amount_of_sleep_hours), # calculating mean
            n = length(amount_of_sleep_hours), # counting the number of observations
            df = n -1, # calculating the degrees of freedom
            sd = sd(amount_of_sleep_hours), # calculating the standard deviation
            se = sd/sqrt(n), # calculating the standard error
            tval = qt(p = 0.05/2, df = df, lower.tail = FALSE), # finding the t value
            margin = tval*se, # calculating the margin of error
            ci_lower = mean - tval*se, # calculating the lower bound of the CI
            ci_upper = mean + tval*se) # calculating the upper bound of the CI
```

```{r jitterplot}
ggplot(data = personal_data_clean, # starting with personal_data_clean data frame
       aes(x = type_of_workout, # x axis is type of workout
           y = amount_of_sleep_hours, # y axis is amount of sleep
           color = type_of_workout)) + # color by type of workout
  geom_jitter(width = 0.1, # changing width of points
              height = 0, # making sure points don't jitter vertically
              alpha = 0.4, # changing transparency
              shape = 21) + # changing shape 
  geom_point(data = personal_data_summary, # plotting the means on the graph
            aes(x = type_of_workout, # x axis is type of workout
                y = mean, # y axis is mean
                color = type_of_workout), # color by type of workout
            size = 2) + # changing size
  geom_errorbar(data = personal_data_summary, # plotting the margins of error
                aes(x = type_of_workout, # x axis is type of workout
                    ymin = mean -se, # setting margins
                    ymax = mean + se,
                    color = type_of_workout), # color by type of workout
                width = 0.1, # changing the width
                inherit.aes = FALSE) + # making sure it changes axis from the first layer
  scale_color_manual(values = c("Aerobic" = "midnightblue", "200 Free" = "mediumvioletred", "500 Free" = "cyan4", "1650 Free" = "steelblue4", "Racing" = "seagreen")) + # changing colors from the default
  labs(x = "Type of Workout", # renaming axis
       y = "Amount of Sleep (hours)",
       title = "Average Sleep and Swimming Workouts", # adding title
       color = "Type of Workout") +
  theme(panel.background = element_blank(), # getting rid of background
        axis.line = element_blank(), # getting rid of lines and ticks
        axis.ticks = element_blank(),
        legend.position = "none") # getting rid of legend


```

## c. Caption

**Figure 1. Type of swimming workout completed compared to average sleep.**
Points represent observations of individual nights of sleep after completing a specific type of swimming workout (total n = 23). Colors represent the different types of workouts (Dark blue = Aerobic, Magenta = 200 Free, Teal = 500 Free, Light blue = 1650 Free, Green = Racing). Bold points represent the mean amount of sleep in hours for each type of workout, and the bars represent the standard error around the mean.
