---
title: "ENVS 193DS Homework 3"
author: "Ellie Schomberg"
date: "05-28-2025"
format: html
---

Link to GitHub repository is [here](https://github.com/ellieschomberg/ENVS-193DS_homework-03.git)

# Set Up

```{r set up}
#| message: FALSE
# reading in packages
library(tidyverse)
library(here)
library(flextable)
library(janitor)
library(readxl)

# reading in data
personal_data <- read.csv("data/updated_personal_data.csv")
```

# Problem 1. Personal Data

## a. Summarizing data

For my project, I am testing to see if the type of swimming workout that I do, each with different yardages and levels of intensity, influences the average amount of sleep I get per night. I am comparing the mean between groups, and I believe that more intense workouts will cause me to sleep more.

## b. Visualization

```{r cleaning data}
personal_data_clean <- read.csv("data/updated_personal_data.csv") |> # start with updated_personal_data data frame
  clean_names() # clean up the column names
```

```{r jitterplot}
ggplot(data = personal_data_clean, # starting with personal_data_clean data frame
       aes(x = type_of_workout, # x axis is type of workout
           y = amount_of_sleep_hours, # y axis is amount of sleep
           color = type_of_workout)) + # color by type of workout
  geom_jitter(width = 0.1, # changing width of points
              height = 0, # making sure points don't jitter vertically
              alpha = 0.4, # changing transparency
              shape = 21) # changing shape

```


